---
title: "Stats 101C project"
author: "Stephen Kim"
date: "November 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(plyr)
library(caret)
library(mice)
library(tree)
library(corrplot)
library(e1071)
library(randomForest)
```


#1.0 Loading Data
```{r}
data = read.csv('HTrainLast.csv', stringsAsFactors = FALSE)
test = read.csv('HTestLastNoY.csv', stringsAsFactors = FALSE)

#Combining training and testing data to clean all data
test$affordabilitty = rep("Affordable", 1500)
data = rbind.data.frame(data,test)
head(data)
data = data[,-1]


numeric_index = which(sapply(data, is.numeric))
num_index = which(sapply(data, is.numeric))
```

#2.0 Feature Engineering

##2.1 Cleaning Missing values

###2.1.1 Total missing Values

```{r}
missing = which(colSums(is.na(data)) > 0)
sort(colSums(sapply(data[missing], is.na)), decreasing = TRUE)
```

First we will eliminate the observations that have a missing value in affordabilitty, which is our response variable
```{r}
which(is.na(data$affordabilitty))
data = data[-c(833,2428),]
```


###2.1.2 Cleaning the predictors Categorical Variables
####Pool QC
```{r}
head(data$PoolQC, n=50)
unique(data$PoolQC)
```
We can see that almost all the observations in PoolQC are missing, this is probably the case because the house does not contain a pool at all therefore we will add another level into PoolQC indicating no pool

```{r}
data$PoolQC[is.na(data$PoolQC)] = "None"
head(data$PoolQC)
```
Now all the NA values are replaced with a level of "None" pool.
But there are some houses that might have a pool area but an NA value in PoolQC

```{r}
data$PoolQC[data$PoolQC == "None" & data$PoolArea != 0]
pool_missing = which(data$PoolQC == "None" & data$PoolArea != 0)
pool_missing
```
Upon further inspection, there are 6 observations that have PoolArea that is not 0 and No pool in PoolQC. I will then give a value into the PoolQC according the the overall house condition to determine a placeholder

```{r}
#Checking the scale of overallcond
unique(data$OverallCond)

#Finding the condition of the houses that have a pool where PoolQC is missing
data$OverallCond[pool_missing]
data$PoolQC[pool_missing] = c("Fa", "Fa", "Fa", "Fa", "Fa", "Fa")
```
Since all the observations had an average house condition, I replaced the "None values with "Fa" which is the "average" of the PoolQC observations

Lastly, because PoolQc is an "ordinal variable" where the order of the levels matter, we want to transform PoolQc into a numerical variable
```{r}
poolvalues = c('None' = 0,'Fa' = 1,'Gd' = 2, 'Ex' = 3)
data$PoolQC = as.integer(revalue(data$PoolQC, poolvalues))
unique(data$PoolQC)
```


####MiscFeature
```{r}
#Observing the unique values in MiscFeature
unique(data$MiscFeature)

data$MiscFeature[is.na(data$MiscFeature)] = "None"
unique(data$MiscFeature)

data$MiscFeature = factor(data$MiscFeature)
```
Because there are not any other columns that provide information for the MiscFeature, we simply we filled all the NA values with None


####Alley
```{r}
unique(data$Alley)
data$Alley[is.na(data$Alley)] = "None"

data$Alley = factor(data$Alley)
unique(data$Alley)
```
Same as the MiscFeature, there are no other columns that provide information for Alley, so we filled NAs with None

####Fence
```{r}
unique(data$Fence)

data$Fence[is.na(data$Fence)] = "None"

data$Fence = factor(data$Fence)
unique(data$Fence)
```
Filled NAs with None

####Fireplace
```{r}
unique(data$FireplaceQu)
```
Looking at fireplace quality, there is fireplaces that can give us information on fireplace quality. First, we will identify which rows do not have a fire place then replace FireplaceQu NA's with None

```{r}
data$FireplaceQu[data$Fireplaces == 0 & !is.na(data$FireplaceQu)]

data$FireplaceQu[is.na(data$FireplaceQu)] = "None"
unique(data$FireplaceQu)
```
There were no inconsistencies within the two columns for fireplaces, so we simply replaced all the NA's with None for FireplaceQu

We also revalue the Fireplace Quality since it is an ordinal variable
```{r}
Quality = c("None" = 0, "Po" = 1, "Fa"= 2, "TA" = 3, "Gd" = 4, "Ex"= 5)

data$FireplaceQu = as.integer(revalue(data$FireplaceQu, Quality))
unique(data$FireplaceQu)
```

####Garage Variables
Since all garage variables have the same missing values, except for one NA, I can just assume that the missing values are houses that do not have a garage.

```{r}
unique(data$GarageType)
data$GarageType[!is.na(data$GarageType)&is.na(data$GarageArea)]

sum(data$GarageType == "Detchd", na.rm = TRUE)
which(!is.na(data$GarageType)&is.na(data$GarageFinish))

#Inspecting the one NA where Garagetype is not missing while others are
data$GarageType[4007]
```


```{r}
#GarageType
unique(data$GarageType)
data$GarageType[is.na(data$GarageType)] = "None"
unique(data$GarageType)

data$GarageType = factor(data$GarageType)

#GarageFinish(Oridnal)
unique(data$GarageFinish)
data$GarageFinish[is.na(data$GarageFinish)] = "None"

finish = c("None" = 0, "RFn" = 2, "Unf" = 1, "Fin"= 3)
data$GarageFinish = as.integer(revalue(data$GarageFinish, finish))
unique(data$GarageFinish)

#GarageQual(Ordinal)
unique(data$GarageQual)
data$GarageQual[is.na(data$GarageQual)] = "None"

data$GarageQual = as.integer(revalue(data$GarageQual, Quality))
unique(data$GarageQual)

#GarageCond(Ordinal)
unique(data$GarageCond)
data$GarageCond[is.na(data$GarageCond)] = "None"

data$GarageCond = as.integer(revalue(data$GarageCond, Quality))
unique(data$GarageCond)
```
After filling in all the NAs, we determined which garage variables seemed ordinal and converted them to integers to utilize their order. Those that were not ordinal was just turned into a factor.


####Bsmt
Looking at the basement variables that have missing values, they all have a similar number of missing values but there are slight differences in the number of missing values

```{r}
data$TotalBsmtSF[is.na(data$BsmtFinType1)]
which(is.na(data$TotalBsmtSF))
```

```{r}
which(is.na(data$BsmtQual) & !is.na(data$BsmtFinType2))
which(is.na(data$BsmtQual) & !is.na(data$BsmtFinType1))
which(is.na(data$BsmtQual) & !is.na(data$BsmtExposure))
miss_bsmt = which(is.na(data$BsmtQual))

data$TotalBsmtSF[miss_bsmt]
miss_bsmt[data$TotalBsmtSF[miss_bsmt] != 0]
```
We can see from this analysis that there are 3 observations where BsmtQuality and BsmtCond is missing but is not missing in BsmtFinType2, BsmtFinType1, and Exposure. There are also values in TotalBsmtSF in these 3 observations therefore we have to input a quality for these observations into BsmtQual and BsmtCond. We will use TotalBsmtArea to estimate BsmtQual and OverallCond to estimate BsmtCond

#####BsmtQual(Ordinal)
```{r}
#BsmtQual
data$TotalBsmtSF[c(39,2598,3358,3513,4220,4947)]/mean(data$TotalBsmtSF, na.rm=T) #these basements are very small compared to the average Basement
data$BsmtQual[c(39,2598, 3358, 3513,4220,4947)] = "Fa"

#Transforming into ordinal
data$BsmtQual = as.integer(revalue(data$BsmtQual, Quality))
unique(data$BsmtQual)
```

#####BsmtCond(Oridnal)
```{r}
#Rows where BsmtCond is missing where others are not
miss_cond=which(is.na(data$BsmtCond) & !is.na(data$BsmtFinType2))
miss_cond

data$OverallCond[miss_cond]

unique(data$BsmtCond)
#Seeing the condition of basements of other houses that have 6 or 7 OverallQual
table(data$BsmtCond[data$OverallCond == 6])
table(data$BsmtCond[data$OverallCond == 7])

#Most are Ta so we will replace these NAs with Ta
data$BsmtCond[miss_cond] = "TA"

#Ordinal transformation
data$BsmtCond = as.integer(revalue(data$BsmtCond, Quality))
unique(data$BsmtCond)

```


#####BsmtExposure(Ordinal)
```{r}
#Observations that are missing in BsmtExposure than aren't in other variables
missing_exp = which(!is.na(data$BsmtQual) & is.na(data$BsmtExposure))
missing_exp

data$OverallQual[missing_exp]

#Trying to come up with an educated guess as to what the Basement exposure would be looking at OverallQual
table(data$BsmtExposure[data$OverallQual == 5])
table(data$BsmtExposure[data$OverallQual == 8])

#We will just assign No exposure to both rows
data$BsmtExposure[missing_exp] = "No"

#Ordinal Transformation
exp_quality = c("No" = 1, "Mn" = 2, "Av" = 3, "Gd" = 4)

data$BsmtExposure = as.integer(revalue(data$BsmtExposure, exp_quality))

unique(data$BsmtExposure)
```
#####BsmtFinType2
```{r}
#There is one observation which is missing where all the other bsmt categoricals are not missing
which(is.na(data$BsmtFinType2) & !is.na(data$BsmtExposure))
unique(data$BsmtFinType2)

data$BsmtUnfSF[2116]
#Since the Unfinish area of the basement seems to be large, we'll just assign this NA as Unf
data$BsmtFinType2[2116] = "Unf"
```


####Assigning None or 0 values to the rest of NAs
```{r}
missing = which(colSums(is.na(data)) > 0)
sort(colSums(sapply(data[missing], is.na)), decreasing = TRUE)
```

```{r}
data$BsmtQual[is.na(data$BsmtQual)] = 0
data$BsmtCond[is.na(data$BsmtCond)] = 0
data$BsmtExposure[is.na(data$BsmtExposure)] = 0
data$BsmtFinType1[is.na(data$BsmtFinType1)] = "None"
data$BsmtFinType2[is.na(data$BsmtFinType2)] = "None"

#Transforming the non ordinal columns as factors
data$BsmtFinType1 = factor(data$BsmtFinType1)
data$BsmtFinType2 = factor(data$BsmtFinType2)
```

####Msn Variables
```{r}
unique(data$MasVnrType)
is.numeric(data$MasVnrArea)
```
There are almost equal number of NAs in between the two variables indicating the NA observations are most likely No masonry area but we will observe the NA which is present in MasVnrType but not in MasVnrArea

```{r}
which(is.na(data$MasVnrType) & !is.na(data$MasVnrArea))
data$MasVnrArea[c(2894,4164,4520)]
table(data$MasVnrType)
data$MasVnrType[c(2894,4164,4520)] = "BrkFace"
unique(data$MasVnrType)
```
Since there is no clear way to determine the specific Mason type, we just replaced it with the most common type
#####MsnVnrType
```{r}
#Now we will fill in the rest of the NAs
data$MasVnrType[is.na(data$MasVnrType)] = "None"
data$MasVnrType = factor(data$MasVnrType)
```

#####MsnVnrArea
```{r}
data$MasVnrArea[is.na(data$MasVnrArea)] = 0
```

####MsZoning
```{r}
unique(data$MSZoning)
table(data$MSZoning)

data$MSZoning[is.na(data$MSZoning)] = "RL"
data$MSZoning = factor(data$MSZoning)
```
We simply replaced the NA values with the mode

####Functional(Ordinal)
```{r}
unique(data$Functional)
table(data$Functional)

data$Functional[is.na(data$Functional)] = "Typ"

#Transforming into ordinal
functional_scale = c('Sal' = 0, 'Sev' = 1, 'Maj2'=3, 'Maj1'=4, 'Mod' =5, 'Min2'=6, 'Min1' = 7,'Typ' =8)

data$Functional = as.numeric(revalue(data$Functional, functional_scale))
```

####Utilties
```{r}
unique(data$Utilities)

data$Utilities[is.na(data$Utilities)] = "AllPub"

data$Utilities = factor(data$Utilities)
```

####Electrical(Ordinal)
```{r}
unique(data$Electrical)
table(data$Electrical)

data$Electrical[is.na(data$Electrical)] = "SBrkr"

#Ordinal Transformation
elec_qual = c('FuseP' = 0, 'FuseF' =1, 'FuseA' = 2, 'SBrkr'= 3)

data$Electrical = as.numeric(revalue(data$Electrical, elec_qual))
unique(data$Electrical)
```


#####KitchenQual(Ordinal)
```{r}
unique(data$KitchenQual)
which(is.na(data$KitchenQual))

data$KitchenAbvGr[c(1458,2735)]
table(data$KitchenAbvGr)

data$KitchenQual = as.numeric(revalue(data$KitchenQual, Quality))
```

####Transforming the rest of the ordinal variables
```{r}
#MSSubClass (not an ordinal variable)
##We will convert this to a categorical variable since it would be better
table(data$MSSubClass)
data$MSSubClass = as.factor(data$MSSubClass)
```
#####Street
```{r}
#Not an ordinal variable, we will simply convert to factor
unique(data$Street)
data$Street = factor(data$Street)
```

#####LotShape
```{r}
#Not an ordinal factor
unique(data$LotShape)
data$LotShape = factor(data$LotShape)
```

#####LandContour
```{r}
#Not an ordinal factor
unique(data$LandContour)
data$LandContour = factor(data$LandContour)
```

#####LotConfig
```{r}
#Not ordinal
data$LotConfig = factor(data$LotConfig)
```

#####LandSlope
```{r}
#Not ordinal
data$LandSlope = factor(data$LandSlope)
```
#####Neighborhood
```{r}
#Not ordinal
data$Neighborhood = factor(data$Neighborhood)
```
#####Condition1&2
```{r}
#Not ordinal
data$Condition1 = factor(data$Condition1)
data$Condition2 = factor(data$Condition2)
```
#####BldgType
```{r}
#Not ordinal
data$BldgType = factor(data$BldgType)
```
#####HouseStyle
```{r}
#Not ordinal
data$HouseStyle = factor(data$HouseStyle)
```
#####RoofStyle
```{r}
#Not ordinal
data$RoofStyle = factor(data$RoofStyle)
```

#####RoofMatl
```{r}
#Not ordinal
unique(data$RoofMatl)
data$RoofMatl = factor(data$RoofMatl)
```
#####Exterior1st&2nd
```{r}
#Not ordinal

#Inspecting the missing values
which(is.na(data$Exterior1st))
which(is.na(data$Exterior2nd))

#Imputing with mode
table(data$Exterior1st)
data$Exterior1st[3870] = "VinylSd"

table(data$Exterior2nd)
data$Exterior2nd[3870] = "VinylSd"

data$Exterior1st = factor(data$Exterior1st)

data$Exterior2nd = factor(data$Exterior2nd)
```

#####ExterQual(ordinal)
```{r}
#ordinal
unique(data$ExterQual)
data$ExterQual = as.numeric(revalue(data$ExterQual, Quality))
```
#####ExterCond
```{r}
#ordinal
data$ExterCond = as.numeric(revalue(data$ExterCond, Quality))
```
#####Foundation
```{r}
#Not ordinal
data$Foundation = factor(data$Foundation)
```
#####Heating
```{r}
#not ordinal
data$Heating = factor(data$Heating)
```
#####HeatingQC(ordinal)
```{r}
#ordinal
data$HeatingQC = as.numeric(revalue(data$HeatingQC, Quality))
```
#####CentralAir
```{r}
data$CentralAir = factor(data$CentralAir)
```

#####PavedDrive
```{r}
unique(data$PavedDrive)
data$PavedDrive = factor(data$PavedDrive)
```

#####SaleType
```{r}
#Not ordinal

#Imputing mode to missing value
table(data$SaleType)
data$SaleType[is.na(data$SaleType)] = "WD"

data$SaleType = factor(data$SaleType)
```

#####SaleCondition
```{r}
data$SaleCondition = factor(data$SaleCondition)
```


####Affordabillity
```{r}
data$affordabilitty = factor(data$affordabilitty)
```

####Rest of Missing Values
We determined that the best way to impute the rest of the values would be with predictive NA imputing, we decided to use the MICE package for that purpose.
```{r}
missing = which(colSums(is.na(data)) > 0)
sort(colSums(sapply(data[missing], is.na)), decreasing = TRUE)

temp <- mice(data, m=5,method="pmm", maxit=15, seed=500)
imputed = complete(temp)
table(is.na(imputed))
```



###3.0 Important variables

####Random forest
#####Using Random forest to see the most important variables
```{r}
head(imputed)
all = imputed
imputed = all[1:3498,]
set.seed(12345)
#Importance plot of all variables
im_var = randomForest(affordabilitty~., data=imputed, ntree=100, importance=TRUE)
varImpPlot(im_var)

#Importance plot of numerical variables
num.tree = randomForest(affordabilitty~., data=imputed[,c(num_index,80)], ntree=100, importance=TRUE)
varImpPlot((num.tree))
```


###4.0 Feature Engineering

With a little research we found that some of the most important variables in determing house price were, size, age and condition, upgrades, neighborhood. Though some of these variables are already in the dataset, we will revisit some and try to create some new variables.

####House size
Here we will combine the sqft area under the ground and above the ground to find the total sqft of the house
```{r}
all$TotalSqFt = all$GrLivArea + all$TotalBsmtSF
```

####House age and condition
Since we already have a variable that determines house condition so we do not need to create it. As for house age, we do not have a variable, so we can make one by using year built and year sold. We also might want to create a variable that indicates whether the house bought was new or not.
```{r}
all$Age = all$YrSold - all$YearBuilt

###Making the "new" variable
new = all$Age <= 1
all$New = factor(new)
```

####Upgrades
For upgrades, the closest variable we would have is YearRemodAdd. These are just the remodel years so we will make a remodeled variable that says yes or no if the house had been remodeled. The year is the same in YearRemodAdd to YearBuilt if the house had not been remodeled
```{r}
all$Remodeled = all$YearRemodAdd != all$YearBuilt
all$Remodeled = factor(all$Remodeled)
```

####Neighborhood
There is already a neighborhood variable so we do not need to create one
```{r}
head(all$Neighborhood)
```

###5.0 Collinearity

```{r}
num_index = which(sapply(all, is.numeric))
num_cor = as.matrix(cor(all[,num_index], use="pairwise.complete.obs"))

high_cor = which(num_cor>0.7 & num_cor!=1, arr.ind=TRUE)#Rows and columns of highly correlated variables
cor_values = num_cor[num_cor>0.7 & num_cor!=1]

cb = cbind(high_cor, cor_values)

cor_data = as.data.frame(cb)
cor_data

names = colnames(as.data.frame(num_cor))
foo = seq(1,51)
names(foo) = names
foo

#Making a column of the col names
cols = vector()
for(i in 1:length(cor_data$col)){
  cols[i] = names(foo)[cor_data$col[i]]
}

cor_data$col = cols
cor_data
```
Now that we have a data frame of all the variables that have high correlations, we will pick the variable of the two variables that seem more important

####Dropping the highly correlated variables
```{r}
drop = c('GarageYrBlt', 'X1stFlrSF', 'TotalBsmtSF', 'GrLivArea', 'Fireplaces', 'GarageArea', 'GarageCond', 'PoolQC', 'ExterQual')
var.out.bool <- !names(all) %in% drop

all = all[,var.out.bool]
```

###6.0 Normalizing and scaling the predictors
```{r}
#Extracting the true numerical predictors from the categorical and the ordinal variables
drop = c('GarageYrBlt', 'X1stFlrSF', 'TotalBsmtSF', 'GrLivArea', 'Fireplaces', 'GarageArea', 'GarageCond', 'PoolQC','MoSold','YrSold','OverallQual','OverallCond', 'Ob','MSSubClass', 'ExterQual')
var.drop = !names(numeric_index) %in% drop
numeric_index[var.drop]

all_numeric = names(all) %in% names(numeric_index[var.drop])

all_num = all[,all_numeric]
all_num$TotalSqFt = all$TotalSqFt
all_num$Age = all$Age

all_cat = all[,!all_numeric]
all_cat = all_cat[,-c(47,48)]
head(all_cat)
```

####Removing skewness in the numerical predictors
```{r}
for(i in 1:ncol(all_num)){
  if(abs(skewness(all_num[,i]))>0.8){
    all_num[,i] = log(all_num[,i]+1)
  }
}

all_num
```
####Scaling the numerical predictors
```{r}
all_num = scale(all_num)
all_num = as.data.frame(all_num)
```

###7.0 One hot encoding the categorical variables
We need to change all the factor variables into numerical because many machine learning algorithms only take in numerical data.
```{r}
#Splitting the categorical data into factors and ordinal
cat_index = which(sapply(all_cat, is.numeric))
all_fac = all_cat[,-cat_index]
all_ord = all_cat[,cat_index]

#Quick fix to YrSold to make into a categorical variable
all_fac$YrSold = factor(all_ord$YrSold)
all_ord = all_ord[,-15]
```

####One hot encoding
```{r}
dmy <- dummyVars(" ~ .", data = all_fac[,-7])
all_dummy <- data.frame(predict(dmy, newdata = all_fac[,-7]))
head(all_dummy)
```


###8.0 Combining the final DataSet
```{r}
final = cbind.data.frame(all_num, all_ord)
final = cbind.data.frame(final, all_dummy)
head(final)
```

```{r}
final_train = final[1:3498,]
final_test = final[3499:4998,]

final_train = final_train[,-c(249,250)]
final_train$affordabilitty = all_cat$affordabilitty[1:3498]
final_test = final_test[,-c(249,250)]
```


###9.0 Models
```{r}
train = sample(1:3498, 2798,replace = F)

model_train = final_train[train,]
model_test = final_train[-train,]
```

###
```{r}
model.tree = randomForest(affordabilitty ~ ., data = model_train, mtry = 30, importance = TRUE, ntree = 300)
varImpPlot(model.tree)
model.tree
preds = predict(model.tree, newdata = model_test[,-258]) 
cm = confusionMatrix(preds, model_test[,258])
fourfoldplot(cm$table)
```


